\documentclass[a4paper,oneside,12pt]{book}

%----------------------------------------------------------------------------------------
%	README!
%   Welcome. It's worth having a read through this file
%   to set up the broad parameters, such as the name of
%   the degree, the school/department, the type of work
%   (dissertation/Final Year Project/report, etc. as well
%   as your own details.
%----------------------------------------------------------------------------------------

%----------------------------------------------------------------------------------------
%	COVER PAGE
%   The cover page is laid out in title/title.tex. You can choose a colour
%   or black and white logo
%----------------------------------------------------------------------------------------

%----------------------------------------------------------------------------------------
%	THESIS INFORMATION
%   Put title, author name, degree, type of work, school, department in here
%   It will be used for the title page and for the embedded PDF information
%----------------------------------------------------------------------------------------
\usepackage{multicol}
\usepackage{float}
\usepackage{booktabs}
\newcommand{\thesistitle}{Comparing Internet Censorship in Ireland vs. Israel} % Your thesis title, this is used in the title and abstract
\newcommand{\degree}{B.A.I. Computer Engineering} % Your degree name, this is used in the title page and abstract
\newcommand{\typeofthesis}{Thesis} % dissertation, Final Year Project, report, etc.
\newcommand{\authorname}{Chris Casey} % Your name, this is used in the title page and PDF stuff
%% Do not put your Student ID in the document, as TCD will not publish
%% documents that contain both your name and your Student ID.

\newcommand{\keywords}{this, that, more} % Keywords for your thesis
\newcommand{\school}{\href{https://www.tcd.ie/scss/}{School of Computer Science and Statistics}} % Your school's name and URL, this is used in the title page

%% Comment out the next line if you don't want a department to appear
\newcommand{\supervisor}{\href{http://www.scss.tcd.ie/}{Dr. Stephen Farrell}} % Your research group's name and URL, this is used in the title page

\AtBeginDocument{
\hypersetup{pdftitle=\thesistitle} % Set the PDF's title to your title
\hypersetup{pdfauthor=\authorname} % Set the PDF's author to your name
\hypersetup{pdfkeywords=\keywords} % Set the PDF's keywords to your keywords
\hypersetup{pdfsubject=\degree} % Set the PDF's keywords to your keywords
}

%% Language and font encodings
\usepackage[T1]{fontenc} 
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{lipsum}
\usepackage{ragged2e} %allows for text alignment preferences

%% Bibliographical stuff
\usepackage[round,sort,comma,numbers]{natbib}

%% Document size
% include showframe as an option if you want to see the boxes
\usepackage[a4paper,top=2.56cm,bottom=2.56cm,left=2.56cm,right=2.56cm, head = 16pt]{geometry}
\setlength{\marginparwidth}{2cm}
%% Useful packages
\usepackage{amsmath}
\usepackage[autostyle=true]{csquotes} % Required to generate language-dependent quotes in the bibliography
\usepackage[pdftex]{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[colorlinks=true, allcolors=black]{hyperref}
\usepackage{xcolor}
\usepackage{caption} % if no caption, no colon
%\usepackage{sfmath} %use sans-serif in the maths sections too
\usepackage[parfill]{parskip}    % Begin paragraphs with an empty line rather than an indent
\usepackage{setspace} % to permit one-and-a-half or double spacing
\usepackage{enumerate} % fancy enumerations like (i) (ii) or (a) (b) and suchlike
\usepackage{booktabs} % To thicken table lines
\usepackage{fancyhdr}

%\pagestyle{plain} % Embrace simplicity!

%% The Mechanical engineers require your name and ID on the top of every page.
%% Uncomment the following block if you want your name and ID at the top of
%% (almost) every page.

\pagestyle{fancy}
\fancyhf{} % sets both header and footer to nothing
\renewcommand{\headrulewidth}{0pt}
\cfoot{\thepage}
%\ifdefined\authorid
%\chead{\it \authorname\ (\authorid)}
%\else
%\chead{\it \authorname}
%\fi
%% End of block

%% It is not a requirement of the university that the font should be sans-serif, but
%% the Mechanical engineers require it. Comment out the following line to disable it
%%\renewcommand{\familydefault}{\sfdefault} %use the sans-serif font as default

%% If you're not using sans-serif, consider using Palatino instead of the LaTeX standard
\usepackage{mathpazo} % Use the Palatino font by default if you prefer it to Computer Modern

\renewcommand{\theequation}{\arabic{equation}} %% use continuous equation numbers

%% Format Chapter headings appropriately
\usepackage{titlesec}
\definecolor{tcdblue}{cmyk}{0.94, 0.38, 0, 0.27}
\newcommand{\hsp}{\hspace{20pt}}
\titleformat{\chapter}[hang]{\Huge\bfseries}{\thechapter\hsp\textcolor{tcdblue}{|}\hsp}{0pt}{\Huge\bfseries}

\title{\thesistitle}
\author{\authorname}

\frontmatter
\begin{document}
\input{title/title.tex}
\pagenumbering{roman}
\doublespacing

\section*{Declaration}
I hereby declare that this \typeofthesis\ is entirely my own work and that it has not been submitted as an exercise for a degree at this or any other university.

I have read and I understand the plagiarism provisions in the General Regulations of the University Calendar for the current year, found at \url{http://www.tcd.ie/calendar}.

I have also completed the Online Tutorial on avoiding plagiarism `Ready Steady Write', located at \url{http://tcd-ie.libguides.com/plagiarism/ready-steady-write}.
\vspace{1cm}

Signed:~\rule{5cm}{0.3pt}\hfill Date:~\rule{5cm}{0.3pt}


\newpage
\chapter{Abstract}

This thesis presents a comparative analysis of internet censorship in Ireland and Israel. Although both are democracies, they differ significantly in political structure and approaches to digital governance. This study investigates how these factors influence the nature, scope, and technical methods of internet censorship in each country. Utilising the Open Observatory of Network Interference (OONI) Probe, both public and curated test suites were used to investigate network interference. The OONI Command Line Interface (CLI) was used to gather ground truth in both Dublin and Tel Aviv.

The tests covered key categories including website accessibility, circumvention tool availability, instant messaging platform functionality, and evidence of middlebox-driven network manipulation. Public OONI data was compared with ground truth in order to derive conclusions about the two countries. EU compliance, the regulation of copyrighted material and the protection of users motivate Ireland's internet censorship. In contrast, Israel has demonstrated its capacity to use internet censorship for political aims.

While both countries show minimal systemic blocking of circumvention tools, isolated cases—such as Psiphon interference were observed. DNS-based blocking was the most common technique in both jurisdictions. This work contributes to the broader understanding of censorship in liberal democracies and highlights the utility of decentralized, open-source measurement tools like OONI Probe in safeguarding global internet freedom.





\newpage
\raggedright %\raggedright turns off justification and hypenation

\section*{\Huge{Acknowledgements}}
I would like to sincerely thank my thesis advisor, Dr. Stephen Farrell, for his invaluable guidance, support, and feedback throughout the course of this research. His insight and encouragement were instrumental in shaping the direction and quality of this work. I would also like to acknowledge Griff Steinman for his contributions as a fellow student and collaborator. His thoughtful input and opinions helped refine several key aspects of the project. I am grateful for the opportunity to work alongside such dedicated and knowledgeable individuals.

\newpage \tableofcontents
%\newpage \listoffigures
%\newpage \listoftables
%\newpage
%\section*{\Huge{Nomenclature}}
%\begin{tabular}{lp{9cm}l}
%A&Area of the wing&$m^{2}$\\
%B\\
%C& Roman letters first, with capitals\ldots\\
%a&then lower case.\\
%b\\
%c\\
%$\Gamma$&Followed by Greek capitals\ldots\\
%$\alpha$&then lower case greek symbols.\\
%$\beta$\\
%$\epsilon$\\
%TLA&Finally, three letter acronyms and other abbreviations arranged alphabetically\\
%\end{tabular}
%\vspace{2cm}

%If a parameter has a typical unit that is used throughout your report, then it should be included here on the right hand side.

%If you have a very mathematical report, then you may wish to divide the nomenclature list into functions and variables, and then sub- and super-scripts.

%Note that Roman mathematical symbols are typically in a serif font in italics.


\mainmatter
\input{introduction/introduction.tex}
\input{Literature Review/Literature Review}
\input{State of the Art/State of the Art}
\input{Circumvention Tools/Circumvention Tools}
\input{Experimental Methods/experiment}
\input{Results/results}
% unsrtnat
\bibliographystyle{unsrtnat}
\bibliography{bibs/sample}

\appendix
\renewcommand{\thechapter}{A\arabic{chapter}}
\input{appendix/appendix.tex}

\end{document}