\chapter{Appendix}

To aid in the writing of this work, additional tools were used to ensure clarity, presentation, correctness, and structure.

\textbf{ChatGPT} ChatGPT was used to assist in structuring the content of this thesis, and in the naming of chapters and sections. It was also used to help generate tables in Latex, and check for spelling and grammar mistakes. ChatGPT also assisted in parsing data collected by the OONI Probe.

\textbf{Grammarly} Grammarly was used to check for spelling and grammar mistakes throughout the entire thesis.

\textbf{Overleaf} Overleaf was used to write and generate the final thesis.

\textbf{GitHub} GitHub was used to store all data related to this project. The public link to this GitHub is \url{https://github.com/ccasey300/censorship_project}.

This work was completed alongside Chris Casey, and some sections of this report were authored by him. These sections include, \ref{sec:Chris-Global-Censorship} and \ref{sec:Chris-EU-Compliance}.

\section{List of Websites}

The list of websites used to test the OONI probe in Ireland and Iraq can be found at the following link on this project's public GibHub:
\url{https://github.com/ccasey300/censorship_project/blob/main/Griff/Files/WebsitesLive.txt}

\section{OONI Data}

All OONI data collected during this work can be found at the following link on this project's public GitHub:
\url{https://github.com/ccasey300/censorship_project/tree/main/Griff/OONIData}

\section{OONI Data Parsing File}

The python script used to parse the OONI data into CSV files can be found at the following link on this project's public GitHub:
\url{https://github.com/ccasey300/censorship_project/blob/main/Griff/Files/OONIDataParse.py}






