\chapter{Circumvention Tools}

Users who care about privacy and anonymity have options to increase their operational security. Some of these tools are outlined below.

\section{Virtual Private Networks}

INSERT general paragraph explaining commercial VPNs, logs policies, etc.

Virtual Private Networks (VPNs) are one of the most commonly used circumvention tools. VPNs work by establishing a private tunnel in the public network through which users traffic is encrypted and routed. This allows users to mask their IP address and change their apparent geo location. This grants both privacy and circumvention potential; VPNs can be used to avoid geo - restrictions by routing traffic through more lenient countries. 

VPNs also protect users from cybercrime through their use of encryption and secure protocols...

The role of VPNs in personal privacy and censorship circumvention cannot be overstated due to how commonplace the technology has become. INSERT A STAT ABOUT VPN USAGE



\section{TOR \& Proxy Solutions}

The Onion Router, originally developed by the US government, is an open-source network overlay that routes internet traffic through volunteer-operated relays. According to the founders "Onion Routing is a distributed overlay network designed to anonymize TCP-based applications like web browsing, secure shell, and instant messaging."\cite{dingledine2004tor}

Requests travel through a relay passing three separate nodes. As a result, it is significantly more difficult to interpret the request’s origin and destination. Tor is also commonly used as a censorship circumvention method. Tor was believed to be secure for a long time but recent developments would suggest otherwise. \cite{tor_not_secure}



\subsection{TOR Bridges}
\subsection{Mesh Networks}
\subsection{SSH Tunneling}


\section{Privacy \& Communication}

\subsection{Encrypted Communication}
\subsection{Email \& Browsing}
\subsection{Tails OS}


