\documentclass[a4paper,oneside,12pt]{book}

%----------------------------------------------------------------------------------------
%	README!
%   Welcome. It's worth having a read through this file
%   to set up the broad parameters, such as the name of
%   the degree, the school/department, the type of work
%   (dissertation/Final Year Project/report, etc. as well
%   as your own details.
%----------------------------------------------------------------------------------------

%----------------------------------------------------------------------------------------
%	COVER PAGE
%   The cover page is laid out in title/title.tex. You can choose a colour
%   or black and white logo
%----------------------------------------------------------------------------------------

%----------------------------------------------------------------------------------------
%	THESIS INFORMATION
%   Put title, author name, degree, type of work, school, department in here
%   It will be used for the title page and for the embedded PDF information
%----------------------------------------------------------------------------------------

\newcommand{\thesistitle}{Comparing Internet Censorship Between Ireland \& Iraq} % Your thesis title, this is used in the title and abstract
\newcommand{\degree}{B.A.I. Computer Engineering} % Your degree name, this is used in the title page and abstract
\newcommand{\typeofthesis}{Thesis} % dissertation, Final Year Project, report, etc.
\newcommand{\authorname}{Griffin Steinman} % Your name, this is used in the title page and PDF stuff
%% Do not put your Student ID in the document, as TCD will not publish
%% documents that contain both your name and your Student ID.

\newcommand{\keywords}{this, that, more} % Keywords for your thesis
\newcommand{\school}{\href{https://www.tcd.ie/scss/}{School of Computer Science and Statistics}} % Your school's name and URL, this is used in the title page

%% Comment out the next line if you don't want a department to appear
\newcommand{\supervisor}{\href{http://www.scss.tcd.ie/}{Dr. Stephen Farrell}} % Your research group's name and URL, this is used in the title page

\AtBeginDocument{
\hypersetup{pdftitle=\thesistitle} % Set the PDF's title to your title
\hypersetup{pdfauthor=\authorname} % Set the PDF's author to your name
\hypersetup{pdfkeywords=\keywords} % Set the PDF's keywords to your keywords
\hypersetup{pdfsubject=\degree} % Set the PDF's keywords to your keywords
}

%% Language and font encodings
\usepackage[T1]{fontenc} 
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{lipsum}
\usepackage{ragged2e} %allows for text alignment preferences

\usepackage{booktabs}
\usepackage{float}
\usepackage{verbatim}

\usepackage{listings}
\usepackage{xcolor}

\lstdefinestyle{python}{
    language=Python,
    basicstyle=\ttfamily\small,
    keywordstyle=\color{blue},
    stringstyle=\color{orange},
    commentstyle=\color{gray},
    showstringspaces=false,
    breaklines=true,
    frame=single,
    tabsize=4,
    captionpos=b
}



%% Bibliographical stuff
\usepackage[round,sort,comma,numbers]{natbib}

%% Document size
% include showframe as an option if you want to see the boxes
\usepackage[a4paper,top=2.56cm,bottom=2.56cm,left=2.56cm,right=2.56cm, head = 16pt]{geometry}
\setlength{\marginparwidth}{2cm}
%% Useful packages
\usepackage{amsmath}
\usepackage[autostyle=true]{csquotes} % Required to generate language-dependent quotes in the bibliography
\usepackage[pdftex]{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[colorlinks=true, allcolors=black]{hyperref}
\usepackage{xcolor}
\usepackage{caption} % if no caption, no colon
%\usepackage{sfmath} %use sans-serif in the maths sections too
\usepackage[parfill]{parskip}    % Begin paragraphs with an empty line rather than an indent
\usepackage{setspace} % to permit one-and-a-half or double spacing
\usepackage{enumerate} % fancy enumerations like (i) (ii) or (a) (b) and suchlike
\usepackage{booktabs} % To thicken table lines
\usepackage{fancyhdr}

%\pagestyle{plain} % Embrace simplicity!

%% The Mechanical engineers require your name and ID on the top of every page.
%% Uncomment the following block if you want your name and ID at the top of
%% (almost) every page.

\pagestyle{fancy}
\fancyhf{} % sets both header and footer to nothing
\renewcommand{\headrulewidth}{0pt}
\cfoot{\thepage}
%\ifdefined\authorid
%\chead{\it \authorname\ (\authorid)}
%\else
%\chead{\it \authorname}
%\fi
%% End of block

%% It is not a requirement of the university that the font should be sans-serif, but
%% the Mechanical engineers require it. Comment out the following line to disable it
%%\renewcommand{\familydefault}{\sfdefault} %use the sans-serif font as default

%% If you're not using sans-serif, consider using Palatino instead of the LaTeX standard
\usepackage{mathpazo} % Use the Palatino font by default if you prefer it to Computer Modern

\renewcommand{\theequation}{\arabic{equation}} %% use continuous equation numbers

%% Format Chapter headings appropriately
\usepackage{titlesec}
\definecolor{tcdblue}{cmyk}{0.94, 0.38, 0, 0.27}
\newcommand{\hsp}{\hspace{20pt}}
\titleformat{\chapter}[hang]{\Huge\bfseries}{\thechapter\hsp\textcolor{tcdblue}{|}\hsp}{0pt}{\Huge\bfseries}

\title{\thesistitle}
\author{\authorname}

\frontmatter
\begin{document}
\input{title/title.tex}
\pagenumbering{roman}
\doublespacing

\section*{Declaration}
I hereby declare that this \typeofthesis\ is entirely my own work and that it has not been submitted as an exercise for a degree at this or any other university.

I have read and I understand the plagiarism provisions in the General Regulations of the University Calendar for the current year, found at \url{http://www.tcd.ie/calendar}.

I have also completed the Online Tutorial on avoiding plagiarism `Ready Steady Write', located at \url{http://tcd-ie.libguides.com/plagiarism/ready-steady-write}.
\vspace{1cm}

Signed:~\rule{5cm}{0.3pt}\hfill Date:~\rule{5cm}{0.3pt}


\newpage
\chapter{Abstract}

This work presents a comparative study of internet censorship in Ireland and Iraq. This work is one in a series of such comparisons. Ireland and Iraq have significantly different political structures, legal frameworks, and cultural norms. As a result, internet censorship is implemented differently in both countries. As a member of the European Union, Ireland implements specifically targeted censorship that calls for legal and regulatory compliance. Iraq demonstrates a more decentralized, reactionary approach, often influenced by political instability and internal events. This work aims to research and analyze each country's scope and technical methods when it comes to internet censorship. The work utilizes the Open Observatory of Network Interference (OONI) Probe to collect network data over a nine-day testing period in both countries. Network testing in Ireland was conducted locally using the OONI Command Line Interface Probe, while network testing in Iraq was done through a Virtual Machine hosted in Baghdad. The data collected included testing for website accessibility, evidence of blocked circumvention tools, instant messaging platform availability, and signs of network manipulation through middleboxes.

Results indicate that both countries exhibit instances of website blocking, with Iraq having a higher block rate of 7\% compared to Ireland's 2.5\%. Blocking methods in both countries include TCP/IP and DNS interference and there is evidence that TLS / SNI-based filtering may be present in some areas of Iraq. Most of Ireland's blocks are related to illegal content, legal cases, and EU compliance (see section \ref{sec:Chris-EU-Compliance}). Iraq's censorship extends to political, religious, and, most often, communication platforms, as internet blackouts are common in the region during times of unrest or national exams. Both countries showed little evidence of systemic efforts to block circumvention tools, but some networks in both Ireland and Iraq showed significant evidence of Psiphon being blocked.

This work aids in the global effort to raise awareness of internet censorship worldwide. Contributing to projects like OONI is beneficial, as it helps researchers identify internet censorship trends across certain countries and raises awareness of network interference. Protecting people's digital rights, privacy, and internet freedoms is growing increasingly important as governments and other entities expand their ability to monitor, restrict, and manipulate online information in ways that can undermine democratic values, personal privacy, and global access to information.




\newpage
\raggedright %\raggedright turns off justification and hypenation

\section*{\Huge{Acknowledgements}}

A special thank you to my mother and father for supporting me.

A special thank you to Eugene O'Rourke and Mark Linnane for their guidance and advice.

\newpage \tableofcontents
%\newpage \listoffigures
%\newpage \listoftables
%\newpage
%\section*{\Huge{Nomenclature}}
%\begin{tabular}{lp{9cm}l}
%A&Area of the wing&$m^{2}$\\
%B\\
%C& Roman letters first, with capitals\ldots\\
%a&then lower case.\\
%b\\
%c\\
%$\Gamma$&Followed by Greek capitals\ldots\\
%$\alpha$&then lower case greek symbols.\\
%$\beta$\\
%$\epsilon$\\
%TLA&Finally, three letter acronyms and other abbreviations arranged alphabetically\\
%\end{tabular}
%\vspace{2cm}

%If a parameter has a typical unit that is used throughout your report, then it should be included here on the right hand side.

%If you have a very mathematical report, then you may wish to divide the nomenclature list into functions and variables, and then sub- and super-scripts.

%Note that Roman mathematical symbols are typically in a serif font in italics.


\mainmatter
\input{introduction/introduction.tex}
\input{Literature Review/Literature Review}
\input{Experimental Methods/experiment}
\input{Results/results}
\input{Security and Privacy/SecurityAndPrivacy}
\input{conclusion/conclusion.tex}

\bibliographystyle{unsrtnat}
\bibliography{bibs/sample}

\appendix
\renewcommand{\thechapter}{A\arabic{chapter}}
\input{appendix/appendix.tex}

\end{document}