\chapter{Security and Privacy}

The following section contains information on the privacy and security
concerns of completing the dissertation. This was completed with an assignment in the CSU44302 Security and Privacy module. In writing a dissertation, it is crucial to consider the potential impacts of the research. This document discusses the security and privacy concerns associated with researching internet censorship.

This section addresses security and privacy concerns involved with operating the OONI probe about this work while considering both the technical and broader legal or regulatory aspects. The comparison of censorship between Ireland and Iraq is significant. While using the OONI probe within both countries comes with its own risks, using the probe in Iraq carries much more concern regarding security and privacy. The environment in Iraq is significantly more dangerous with ongoing government surveillance, frequent shutting down of social media sites, and the risk of authorities considering unauthorized data-gathering activities as suspicious. All these factors indicate the need to carefully plan where, how, and why measurements are taken and how resulting data will be stored.

\section{OONI}

Although OONI strives to minimize the collection of personal data, its measurements are published openly, which may inadvertently disclose approximate locations and times when tests occurred \cite{ooniOONIPrivacyAndSecurity}. If the individual running OONI is tied to a VM in Iraq with an IP address, local authorities or ISPs might link test activity back to the source. This risk is particularly heightened if the probe frequently connects to or tests politically sensitive, banned, or controversial websites. In a high-censorship environment, repeated network tests can attract attention and might be interpreted as an intentional challenge to government policies.

(Chris paragraph below)

The claim emphasizes their positive track record: “To our knowledge, no OONI Probe user has ever faced consequences as a result of using our software.”\cite{OONIRisks}. The success of OONI is critically dependent on users conducting tests without repercussions. However, OONI outlines several scenarios where running their probe may be unwise. This includes users residing in countries with a history of prosecuting similar activities, surveillance concerns, or legal restrictions on accessing content. Users in one or more categories should be wary of the potential
risks. In this context, operating in Ireland with no reason to believe I am under surveillance, I am considered a low-risk user.

\section{Iraq Virtual Machine}

When deploying a VM in Iraq, the potential security and privacy risk increases because authorities or other outside sources might attempt to compromise the server. The government of Iraq might be motivated to confiscate or check the contents of the VM to identify individuals who are actively monitoring sensitive network interference. It is also possible that the hosting provider itself can be forced to give logs, user connections, or site testing targets, which removes all privacy the user has. To avoid this, one should ensure that no personal data is used on this VM, and tools for circumvention are used to encrypt the origin of the user accessing the VM.

Even beyond direct government intervention, third-party hacks or malware injection is risky. A public and well-known measurement platform like OONI will attract hackers looking to disrupt users of this tool or introduce malware that will capture all incoming and outgoing traffic. In Iraq, the network infrastructure might already contain middleboxes or deep packet inspection systems actively filtering or manipulating data. These devices sometimes disrupt the traffic generated by the OONI probe measurements, leading to manipulated data being collected.




